Keystoneness, centrality, and the structural controllability of ecological networks

### 2019-01-14 paper/reviewers-response.Rmd

> "The methods are mostly clear but there is no description of how the 'likelihood of being a superior node' is calculated."

<!-- DBS: Don't you think they're asking for this to appear in the main text and not the supps? -->
*Accepted and changed*. Thanks for pointing this out. We have modified the sentence mentioning 'the likelihood of being a superior node' in the first paragraph of the section \@ref(S-reciprocal-links) to "...and whether it is a superior node..." to avoid confusion. Then we define the 'probability of being a superior node' at the end of the section: "It is important to note that a species that is found to be a superior node under a particular control configuration will be so in all possible control configurations for the particular non-reciprocal network. 
We therefore denote the average of $\sigma$ across all non-reciprocal versions as the probability of the species being a superior node in its community".

> "I would suggest including clarification on the limitations of using control capacity to summarise species' importance, and examples to help the reader build an intuition regarding the control capacity metric. As pointed out, networks that give rise to a species having a high control capacity can be quite distinct to networks that give rise to a species having a high value of particular centrality metrics.  Providing an example in the Discussion of, for instance, those species that have a control capacity of 1 but a low degree would be very useful."

*Partially accepted and changed*: We think that more studies are needed before we can accurately build an intuition of what constitutes a species with a high control capacity, especially in regards to other kinds of networks. 
As such we have added a clarification in \lr{M-R2-clarification-limitations} that states "As much potential as any metric or metrics to summarise species' importance might appear to have, it's clear that we also need more empirical studies in different types of networks in order to build intuition and ground truth their usefulness". 
We also considered the suggestion about including a specific example, but we ultimately concluded that none would be particularly didactic for a reader unfamiliar with the exact empirical context.

### 2019-01-14 paper/reviewers-response.Rmd

> "*L430*: 'marginally affected by the topology of their interactions'"

*Accepted and changed*: We have rephrased this line as: "By accounting for network dynamics (even if in a simple way), structural stability incorporates more ecological realism, especially in the extreme scenario in which the structure of interactions within the community only marginally affects the community's state" (\lr{M-R2-marginally-affected}).

> "*L436* and *L441*: 'dominate the state [of their community]'"

*Accepted and changed*: Because we don't have a proper definition, we have modified the conclusion paragraph and now the affected sentences read: "Collectively, critical species have the ability to strongly influence the state of their community and therefore are likely to be highly relevant for ecosystem management and conservation. 
While useful, centrality metrics---which have often been used as a proxy for keystoneness---fail to identify some of these species, highlighting their limitations when we fully embrace the notion that ecological communities are dynamical systems.
Paine (1969) showed nearly 50 years ago that a single species can sometimes fill this role on its own" (\lr{M-R2-dominate}).

### 2019-01-14 paper/manuscript.Rmd
Add sentence about limitations of controllability
Remove 'dominate' from conclusion as we dont have a nice definition

Other conceptual differences between structural controllability and centrality metrics provide three key insights into the conservation of ecological networks. 
\R{R2-structural-controllability} First, structural controllability emphasizes that the effect a species has on other species is not independent of the effects that the other species have in the community. 
The rankings provided by centrality metrics and other heuristics fail to account for the collective influence of several species at once. 
Second, it demonstrates that to ensure the persistence of a community it is often necessary to consider the abundances of more than a single species, even when full control is infeasible or undesired (for example `r round(sum(crit$n_crit > 1)/nrow(crit)*100)`% of our communities contained more than one critical species). 
Third, structural controllability explicitly acknowledges the existence of multiple management strategies and some will be better than others depending on the context.
Approaches to prioritise species for conservation and reintroduction based on traits or centrality are still useful and are likely to overlap with species' control capacity [@pires_friendship_2017; @devoto_understanding_2012]. 
Stepping back, our results also provide support to the idea that management decisions should not be based on a single technique but indicate that focusing on ecosystem processes and interactions may be more effective than traditional ranking-based approaches [@harvey_bridging_2017].

@@

Here we show that structural controllability can be applied in an ecological setting to gain insight into the stability of a community and the role that species play in modifying the abundance of other species and ultimately the state of the community. 
These characteristics make structural stability an ideal framework to evaluate the effects of invasions and other types of perturbations. 
Importantly, structural controllability can be used to identify critical species in the community that promote biodiversity and underpin the stable coexistence of species in their community.
Collectively, critical species have the ability to strongly influence the state of their community and therefore are likely to be highly relevant for ecosystem management and conservation. 
While useful, centrality metrics---which have often been used as a proxy for keystoneness---fail to identify some of these species, highlighting their limitations when we fully embrace the notion that ecological communities are dynamical systems.
@paine_note_1969 showed nearly 50 years ago that a single species can sometimes fill this role on its own. 
Structural controllability suggests that this situation might be the exception rather than the rule. 
We see our study as a starting point to study the controllability of ecological and socio-ecological systems where many exciting questions lie ahead.

>>>>

Other conceptual differences between structural controllability and centrality metrics provide three key insights into the conservation of ecological networks. 
\R{R2-structural-controllability} First, structural controllability emphasizes that the effect a species has on other species is not independent of the effects that the other species have in the community. 
The rankings provided by centrality metrics and other heuristics fail to account for the collective influence of several species at once. 
Second, it demonstrates that to ensure the persistence of a community it is often necessary to consider the abundances of more than a single species, even when full control is infeasible or undesired (for example `r round(sum(crit$n_crit > 1)/nrow(crit)*100)`% of our communities contained more than one critical species). 
Third, structural controllability explicitly acknowledges the existence of multiple management strategies and some will be better than others depending on the context.
Approaches to prioritise species for conservation and reintroduction based on traits or centrality are still useful and are likely to overlap with species' control capacity [@pires_friendship_2017; @devoto_understanding_2012]. 
Stepping back, our results also provide support to the idea that management decisions should not be based on a single technique but indicate that focusing on ecosystem processes and interactions may be more effective than traditional ranking-based approaches [@harvey_bridging_2017]. 
\R{R2-clarification-limitations}As much potential as any metric or metrics to summarise species' importance might appear to have, it's clear that we also need more empirical studies in different types of networks in order to build intuition and ground truth their usefulness.

@@

Here we show that structural controllability can be applied in an ecological setting to gain insight into the stability of a community and the role that species play in modifying the abundance of other species and ultimately the state of the community. 
These characteristics make structural stability an ideal framework to evaluate the effects of invasions and other types of perturbations. 
Importantly, structural controllability can be used to identify critical species in the community that promote biodiversity and underpin the stable coexistence of species in their community.
\R{R2-dominate}Collectively, critical species have the ability to strongly influence the state of their community and therefore are likely to be highly relevant for ecosystem management and conservation. 
While useful, centrality metrics---which have often been used as a proxy for keystoneness---fail to identify some of these species, highlighting their limitations when we fully embrace the notion that ecological communities are dynamical systems.
@paine_note_1969 showed nearly 50 years ago that a single species can sometimes fill this role on its own. 
Structural controllability suggests that this situation might be the exception rather than the rule. 
We see our study as a starting point to study the controllability of ecological and socio-ecological systems where many exciting questions lie ahead.
